---
title: "Write Down All The Things"
author: "Kui Xue"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    number_sections: yes #是否自动按节编号
    toc: yes # table of content
   
 
---

```{r include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(warning=FALSE,message=FALSE,collapse = TRUE, fig.align="center",fig.width = 8)
```

<!-- include= FALSE, hide code chunk above. -->

> “I don’t need to write that down, I’ll remember it” is the biggest lie you can tell yourself! Write down everything you do — even if it doesn’t work. This includes meeting notes, method details, code annotations, among other things.






## Statistic

[**Repeated measures ANOVA**](https://www.datanovia.com/en/lessons/repeated-measures-anova-in-r/)

判断被试内因子是否存在交互作用，首先看是否满足球形假设p>.05。若不满足，需校正。

**One-way repeated measures ANOVA**

within-subjects factor (here time)

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
data('selfesteem', package = 'datarium')
head(selfesteem, 3)

self1= selfesteem %>% 
  pivot_longer(-1, names_to = 'time', values_to = 'score')
head(self1, 3)
```
```{r}
# assumption 1: identify outliers
self1 %>% 
  group_by(time) %>% 
  identify_outliers(score)

# assumption 2: normality
self1 %>% 
  group_by(time) %>% 
  shapiro_test(score)
```


<div id="yellowy" style="color:#c19954; border:solid 1px #fbeccd; border-radius:5px; background-color: #fdf9e4; margin-right: 2cm; padding:10px 10px 5px 15px;">

*Note that*, if your sample size is greater than 50, the normal QQ plot is preferred because at larger sample sizes the Shapiro-Wilk test becomes very sensitive even to a minor deviation from normality.
</div>


```{r}
# assumption 3: sphericity
anova_test(self1, dv= score, wid= id, within= time)
```

```{r}
# pairwise comparisons; post hoc test
self1 %>%
  pairwise_t_test(score ~ time, paired = TRUE,
    p.adjust.method = "bonferroni")
```





**Two-way repeated measures ANOVA**

<div id="blue" style="border-top:solid 2px #366393; background-color: #f2f8ff; margin-right: 2cm; padding:10px 10px 5px 15px;">
The main goal of two-way and three-way repeated measures ANOVA is, respectively, to evaluate if there is a statistically significant interaction effect between two and three within-subjects factors in explaining a continuous outcome variable.
</div>

```{r}
data('selfesteem2', package = 'datarium')
set.seed(10)
selfesteem2 %>% sample_n_by(treatment, size= 2)

self2= selfesteem2 %>% 
  pivot_longer(-1:-2, names_to = 'time', values_to = 'score')
```

Check assumptions for outliers and normality using vars 'treatment' and 'time' to group by

```{r}
self2 %>% 
  ggqqplot('score')+
  facet_grid(vars(treatment), vars(time))+
  theme_bw()
```

```{r}
self2 %>% 
  anova_test(dv= score, wid= id, within= c(treatment, time))
```


<div id="green" style="color:#478948; border:solid 1px #d7eac7; border-radius:5px; background-color: #e0f1d9; margin-right: 2cm; padding:10px 10px 5px 15px;">

There is a significant interaction (p= 4.63e-07), indicates that effect of one factor 
(e.g., treatment) on y depends on the level of the other factor (e.g., time).

In the situation where we can separate two-way interaction into:

- **Simple main effect**
- **Simple pairwise comparison**: if simple main effect is significant, run pairwise comparisons to determine which groups are different.


</div>


**Procedure for a significant two-way interaction**

```{r eval=FALSE}
# Effect of treatment at each time point
self2 %>% 
  group_by(time) %>% 
  anova_test(dv= score, wid= id, within= treatment) %>% 
  adjust_pvalue(method = 'bonferroni')

self2 %>%
  group_by(time) %>%
  pairwise_t_test(score ~ treatment, paired = TRUE,
    p.adjust.method = "bonferroni")
```

<div id="yellowy" style="color:#c19954; border:solid 1px #fbeccd; border-radius:5px; background-color: #fdf9e4; margin-right: 2cm; padding:10px 10px 5px 15px;">

Note that, the treatment factor variable has only two levels (“ctr” and “Diet”); thus, ANOVA test and paired t-test will give the same p-values.
</div>

**Procedure for non-significant two-way interaction**

If the interaction is not significant, you need to interpret the main effects for each of the two variables: treatment and time.

```{r eval=FALSE}
# comparisons for treatment variable
self2 %>%
  pairwise_t_test(score ~ treatment, paired = TRUE, 
    p.adjust.method = "bonferroni")
# comparisons for time variable
self2 %>%
  pairwise_t_test(score ~ time, paired = TRUE, 
    p.adjust.method = "bonferroni")
```

**Three-way repeated measures ANOVA**

```{r results='hide'}
data('weightloss', package = 'datarium')
w1= weightloss %>% 
  pivot_longer(-id:-exercises, names_to = 'time', values_to = 'score')

w1 %>% 
  anova_test(dv= score, wid= id, within = c(diet, exercises, time))
```

Note that, if the three-way interaction is not statistically significant, you need to consult the two-way interactions in the output.

Considering there is a significant three-way interaction, we execute:

- Simple two-way interaction, run two-way interaction at each level of third var
- Simple main effect, run one-way model at each level of second var, and
- pairwise comparison, if necessary

You are free to decide which two variables will form the simple two-way interactions and which variable will act as the third (moderator) variable. 


```{r}
# post-hoc test
w1 %>% 
  group_by(diet) %>% 
  anova_test(dv= score, wid= id, within= c(exercises, time)) %>% 
  get_anova_table()
```


<div id="yellowy" style="color:#c19954; border:solid 1px #fbeccd; border-radius:5px; background-color: #fdf9e4; margin-right: 2cm; padding:10px 10px 5px 15px;">

Note that, it’s recommended to adjust the p-value for multiple testing. One common approach is to apply a Bonferroni adjustment to downgrade the level at which you declare statistical significance.

This can be done by dividing the current level you declare statistical significance at (i.e., p < 0.05) by `the number of simple two-way interaction you are computing` (i.e., 2).

Thus, you only declare a two-way interaction as statistically significant when p < 0.025 (i.e., p < 0.05/2). Applying this to our current example, we would still make the same conclusions.
</div>

A statistically significant simple two-way interaction can be followed up with simple simple main effects.

We will only need to consider the result of the simple simple main effect analyses for the `“diet no” trial` as this was the only simple two-way interaction that was statistically significant

```{r}
w1 %>% 
  group_by(diet, exercises) %>% 
  anova_test(dv= score, wid= id, within= time) %>% 
  get_anova_table() %>% 
  filter(diet== "no")  # interaction only existing in 'diet no' trial
```


```{r}
# compute simple pairwise comparison
w1 %>% 
  filter(diet== "no" & exercises== "yes") %>% 
  pairwise_t_test(score ~ time, paired = TRUE, p.adjust.method = "bonferroni")

# comparison for 'diet yes' trial; non-significant interaction
w1 %>% 
  filter(diet== "yes") %>% 
  pairwise_t_test(score ~ time, paired = TRUE, p.adjust.method = 'bonferroni')

```







   

